---
title: "Hands-on_Ex02"
author: "Gao Ya"
date: "January 17, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
---

# 1 Load packages

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 

```

# 2 Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# 3 Beyond ggplot2

## 3.1 Beyond ggplot2 Annotation: ggrepel

ggrepel provides geoms for ggplot2 to repel overlapping text labels:

`geom_text_repel()`, `geom_label_repel()`

::: panel-tabset
## With ggrepel

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point(aes(color = ENGLISH+MATHS)) +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## Without ggrepel

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point(aes(color = ENGLISH+MATHS)) +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

## 3.2 Beyond ggplot2: Themes

ggplot2 comes with eight built-in themes, they are: `theme_gray()`, `theme_bw()`, `theme_classic()`, `theme_dark()`, `theme_light()`, `theme_linedraw()`, `theme_minimal()`, `and theme_void()`.

These are complete themes which control all non-data display. Use `theme()` if you just need to tweak the display of an existing theme.

::: panel-tabset
## theme_classic

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey90", 
                 fill="cadetblue3") +
  theme_classic() +
  ggtitle("Distribution of Maths scores") 

```

## theme_dark

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey90", 
                 fill="antiquewhite3") +
  theme_dark() +
  ggtitle("Distribution of Maths scores") 

```

## theme_minimal

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey90", 
                 fill="azure3") +
  theme_minimal() +
  ggtitle("Distribution of Maths scores") 

```
:::

### 3.2.1 Working with ggtheme package

ggthemes provides ‘ggplot2’ themes that replicate the look of plots by Edward Tufte, Stephen Few, Fivethirtyeight, The Economist, ‘Stata’, ‘Excel’, and The Wall Street Journal, among others.

::: panel-tabset
## theme_economist

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()

```

## theme_fivethirtyeight

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_fivethirtyeight()

```
:::

### 3.2.2 Working with hrbthems package

hrbrthemes package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 10,
              base_size = 10,
              grid = "Y")
```

## 3.3 Beyond Single Graph

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs.

### 3.3.1 Creating Composite Graphics

We can plot more complex composite by using appropriate operators. For example, the composite figure below is plotted by using:

“\|” operator to stack two ggplot2 graphs, “/” operator to place the plots beside each other, “()” operator the define the sequence of the plotting.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
sct <- ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point(aes(color = ENGLISH+MATHS)) +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores") + theme(text=element_text(size=4), 
        axis.text=element_text(size=4),
        axis.title=element_text(size=5),
        plot.title=element_text(size=7), 
        legend.text=element_text(size=4),
        legend.title=element_text(size=4))

box <- ggplot(exam_data, aes(x=GENDER, y=ENGLISH+MATHS)) +
  ggtitle("Gender distribution of Maths +English scores")+ 
  geom_boxplot()+ theme(text=element_text(size=4), 
        axis.text=element_text(size=4),
        axis.title=element_text(size=5),
        plot.title=element_text(size=7), 
        legend.text=element_text(size=4),
        legend.title=element_text(size=4))


hist <- ggplot(data=exam_data, 
             aes(x = MATHS+ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey90", 
                 fill="azure3") +
  theme_minimal() +
  ggtitle("Distribution of Maths +English scores")+ theme(text=element_text(size=4), 
        axis.text=element_text(size=4),
        axis.title=element_text(size=5),
        plot.title=element_text(size=15), 
        legend.text=element_text(size=4),
        legend.title=element_text(size=4))

(sct | box ) /
      hist
```

### 3.3.2 Creating figure with insert

```{r}
#| code-fold: true
#| code-summary: "Show the code"
hist + inset_element(box, 
                   left = 0.02, 
                   bottom = 0.3, 
                   right = 0.5, 
                   top = 1)
```

### 3.3.3 Creating a composite figure by using patchwork and ggtheme

```{r}
#| code-fold: true
#| code-summary: "Show the code"
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")


p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")


p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")


patchwork <- (p1 / p2) | p3
patchwork & theme_economist()+ theme(
        plot.title=element_text(size=12))
```

# 4 References

[ggrepel](https://ggrepel.slowkow.com/)

[ggthemes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

[Built-in color names in R](http://www.sthda.com/english/wiki/colors-in-r)

[patchwork](https://patchwork.data-imaginist.com/)
